---
layout: review, 11pt
linenumbers: true
title: "rOpenSci tools for textmining open source science literature"
author:
  - name: Scott Chamberlain
    affiliation: cstar
    email: scott(at)ropensci.org
    footnote: Corresponding author
address:
  - code: cstar
    address: |
      rOpenSci, Museum of Paleontology, University of California, Berkeley, CA, USA
abstract: |
      Corresponding Author:

      Scott Chamberlain

      rOpenSci, Museum of Paleontology, University of California, Berkeley, CA, USA

      Email address: [scott@ropensci.org](mailto:scott@ropensci.org)

      \newpage

      Background. xxxx.

      Methods. xxxx.

      Results. xxxx.

      Discussion. xxxx.

bibliography: components/references.bib
csl: components/peerj.csl
documentclass: components/elsarticle

output:
  pdf_document:
    template: components/elsarticle.latex
    keep_tex: true
    fig_caption: true
---


```{r compile-settings, include=FALSE}
library("methods")
library("knitr")
opts_chunk$set(
  tidy = FALSE,
  warning = FALSE,
  message = FALSE,
  cache = 1,
  comment = '#>',
  collapse = TRUE,
  verbose = TRUE
)

basename <- gsub(".Rmd", "", knitr:::knit_concord$get('infile'))
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))

# tibble options
options(tibble.max_extra_cols = 10)
```

\newpage

# Introduction

Explosion of digital context online. But we need tools to facilitate textmining (i.e., rOpenSci)

Why text mine it.

How to text mine it.

Here, we introduce and give examples for textmining in R using rOpenSci packages, and related packages.


# Text mining sources

[Describe open source data for text mining that we link to in R, and those we don't link to in R.]

There is increasing open source scientific literature content available online. However, only a small proportion of scientific journals provide access to their full content; whereas, many publishers provide open access to their metadata only (Table 1).

#### Table 1. Sources of scientific literature, their content type provided via web services, whether rOpenSci has an R packages for the service, and where to find the API documentation.

Data Provider                      | Content Type            | rOpenSci? | API Documentation
---------------------------------- | ---------------------   | --------- | -----------------
Public Library of Science (PLoS)   | Full text/altmetrics    | rplos     | http://api.plos.org/
Springer                           | Full text on OA content | rspringer | http://dev.springer.com/
Pensoft                            | Full text/altmetrics    | rpensoft  | http://bit.ly/KYP0Zi
Nature Publishing Group            | Metadata only           | No        | http://developers.nature.com/
Mendeley                           | Metadata only           | rmendeley | http://dev.mendeley.com/
DataCite                           | Metadata only           | rdatacite | http://oai.datacite.org/
Biodiversity Heritage Library      | Full content            | rbhl      | http://bit.ly/KYQ1Rd
Scopus (Elsevier)                  | Metadata only ????      | No        | http://bit.ly/J9S616

#### Table 2. Tools for text mining science literature.

Tool           |  Platform            | Services              | URL
-------------- | -------------------- | --------------------- | ---------------------
PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
rOpenSci       | R                    | See Table 1           | http://ropensci.org/
PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY
PMC Miner      | Python               | PubMed                | http://bit.ly/L29ekY


# How to text mine from R: Three case studies

### Case study 1

```{r city-diversity, child="use-cases/city-diversity.Rmd", eval=FALSE}
```

### Case study 2

```{r city-diversity, child="use-cases/city-diversity.Rmd", eval=FALSE}
```

### Case study 3

```{r city-diversity, child="use-cases/city-diversity.Rmd", eval=FALSE}
```

# Conclusions and future directions



# Acknowledgments

# Data Accessibility

All scripts and data used in this paper can be found in the permanent
data archive Zenodo under the digital object identifier (DOI). This DOI
corresponds to a snapshot of the GitHub repository at
<https://github.com/ropensci/textmine>. Software can be found at
<https://github.com/ropensci/xxx>, xxxx, all under MIT licenses.


# References
